Galaxy Zoo: Morphological Classifications for 120,000 Galaxies in HST Legacy Imaging

### 2016-05-29 writeup/gz_hubble_data.tex
Funding source for TM, re #69.
various typos corrected; changed bounding box for figure 8

The technique for redshift debiasing (Section~\ref{sec:debiasing}) requires
consistent measurements of the galaxy surface brightness. For both the
\redshifted{} calibration images and the \hst{} images which have their
morphologies corrected, we calculate the mean surface brightness $\mu$ within the
:w
effective radius ($R_e$) as:

@@

Figure~\ref{fig:zeta_examples} shows the change in vote fraction and the
best-fit model for a random set of galaxies in the \ferengi{} sample. The
results show that there is a clear dependence on redshift for the observed
changes in \ffeatures. To examine whether a better fit should involve
additional parameters, we examine the global dependence of $\zeta$ on the
galaxy surface brightness. This is motivated by the fact that brighter galaxies
would presumably be easier to identify; \citet{bam09} and \citet{wil13} found
changes in \ffeatures{} depending on a combination of physical size and
absolute magnitude. To simplify our model, we subsume these into a single
parameter $\mu$.

@@

\begin{figure}
\center
\includegraphics[width=0.50\textwidth]{figures/zeta_mu.pdf}
\caption{All fits for the \ferengi{} galaxies of the vote fraction dropoff
parameter $\zeta$ for \ffeatures{} as a function of surface brightness. This
includes only galaxies with a bounded range on the dropoff
($-10<\log(\zeta)<10$) and sufficient points to fit each function (37~original
galaxies, each artificially redshifted in eight bins over a range from $0.3\lesssim z_\mathrm{sim}\lesssim1.0$).}
\label{fig:zeta_mu}
\end{figure}

@@

The criteria for determining whether a region of this space is single-valued,
and therefore correctable, is as follows: In each surface brightness and
redshift bin, the relationship between \ffeaturesz{} and \ffeaturesrest{} is
modeled by fitting the data with polynomials of degrees $n=3$, 2, and 1, and
using the best formal fit out of the three as measured by the sum of the
residuals. These fits are shown as the dashed black lines in
Figure~\ref{fig:f_vs_f}. Flat regions of the bins are areas in which there is
not a clear single-valued relationship between \ffeaturesz{} and
\ffeaturesrest. We quantify this by measuring the slope of the best-fit
polynomical; regions of the bins with a slope less than 0.4 are considered
\emph{not} one-to-one, and therefore \ffeaturesz{} cannot be boosted to its
\ffeaturesrest{} value.  Galaxies in this region are referred to as the
\emph{lower limit} sample, because the most stringent correction available is
that the weighted \ffeatures{} is a lower limit to the true value.  These
regions are highlighted in blue in Figure~\ref{fig:f_vs_f}. Uncolored (white)
regions of the plot have sufficiently high slopes to consider the relationship
as single-valued; galaxies in these regions are considered ``correctable'', and
only these are used in measuring the parameters for the $\zeta$ function
(Section~\ref{ssec:zeta}). Only surface brightness/redshift bins with at least
5~galaxies were considered; regions with fewer than 5~galaxies are considered
to have ``not enough information'' to determine the \ffeaturesz{} and
\ffeaturesrest{} relationship, colored gray in Figure~\ref{fig:f_vs_f}. These
galaxies are hereafter referred to as the ``NEI'' sample.

@@

For the ``lower-limit'' galaxies for which a single debiased \ffeatures{} value
cannot be confidently assigned, the \emph{range} of debiased values is
estimated using a method visualized in Figure~\ref{fig:f_vs_f_zoom}. This uses the
\ferengi{} simulated data to analyze the range of intrinsic \ffeaturesrest{}
values for any given observed \ffeatures{} value, again as a function of
surface brightness and redshift. In each unclassifiable (shaded blue) $z$,$\mu$ bin, the spread of
intrinsic values of \ffeaturesrest{} for four ranges of observed \ffeatures is computed.
The range of intrinsic values is defined as the inner 80\% of the data,
represented by the orange bars in Figure~\ref{fig:f_vs_f_zoom}. For any galaxy which
cannot be directly debiased by the $\zeta$ method, these ranges are used to
denote the upper and lower limits on the expected values \ffeaturesrest{} as a
function of the observed \ffeatures.

@@

Few galaxies in the sample have sufficiently high corrections to completely
change them from being confidently ``smooth'' to ``featured'' following the
correction for redshift bias (Figure~\ref{fig:debiased_corrections}). As a
check, highly boosted galaxies are compared to the expert classifications in
CANDELS \citep{kar15}.  There are only eight galaxies that were strongly
boosted (\ffeatures$<0.2$ and \fbest$>0.5$) in GZH and also appears in the
CANDELS expert sample. Of those eight galaxies, \citet{kar15} classify 5 as
spheroids, 2 as disks, and 1 as irregular/disky. The \fbest{} values for GZH
are all between 0.5 and 0.6, making them intermediate disk candidates. One of
the spheroids has a hint of extended emission to the south that may have been
missed by the CANDELS experts, but almost all images are dim and with very low
surface brightnesses in the ACS imaging, making positive identification
challenging. Since the surveys use different rest-frame filters and the overlap
sample is tiny, though, detailed comparisons between the overall morphologies
are highly difficult.

@@

Each image is listed under a unique project ID (eg,~AHZ000001); the actual
galaxy in the image is identified by the combination of the OBJNO and original
survey. For each of the 55~responses in the GZH decision tree, the following
classification data is provided: for each question, $\tt N_{votes}$ is the
number of classifiers who answered that question. For each unique answer, $\tt fraction$
is the fraction of classifiers who selected that answer ($\rm N_{answer}/N_{votes}$, and
$\tt weighted$ is the weighted fraction, which takes into account overall
consistency (Section~\ref{ssec:weighting}).

@@

The comparison and analysis of the GZH morphologies is done a set of matched,
volume-limited samples at $z<1.0$ and $m_{I|i|z} < 22.5$~mag.  The redshift and
magnitude limits for the sample are chosen to match the shared constraints for
the shallowest depths \citep[GEMS;][]{bun05} and the limits on morphological
reliability \citep[COSMOS/ZEST;][]{sca07} in the comparison catalogs. We match
all catalogs against GZH using a positional radius of 0.5\arcsec.

@@

Figure~\ref{fig:comparisons_odd} analyzes the same sets of matched galaxies, but
compares the data the \fodd{} parameter in GZH to examine its correlation with
irregular morphologies. The relationship in this case is less clear; for the
COSMOS catalogs, in particular, early-types, spirals, and irregulars are
roughly equally probable at \fodd$>0.5$. In both AEGIS and GOODS (there is no
measurment of irregular structure for GEMS in \citealt{hau07}), the proportion
of irregular galaxies does increase with \fodd. However, the actual number of
galaxies at \fodd$>0.5$ is quite small ($N\lesssim100$) in both samples, and so
estimating an accurate fraction is less precise.  We also note that this
particular question ({\it ``anything odd?''}) has somewhat uncertain phrasing
in GZH, and that the design of the tree explicitly separates clumpy from
irregular or merging morphologies, which would have similar properties in
automatic techniques. A more accurate selection of irregular morphologies in
GZH should use a weighted combination of \fbest, \fclumpy, and \fodd.

@@

To determine the rate at which the vote fraction for any Task decreases with
redshift or surface brightness, data from simulated \ferengi{} images were
modeled with functions in discrete redshift and surface brightness bins
(Section~\ref{ssec:zeta_results}). A cut of $N>10$ was placed on the number of
votes to reduce the error in the vote fractions and ensure they were
well-sampled before applying a fit to the data. This requirement was only met
for Task~01, which significantly reduced the goodness-of-fit for the model
functions as compared with to the smooth/features task. Here we show an
example of results obtained for the bar Task. Figure~\ref{fig:f_vs_fbar}
shows that both the low sample rate and the degree of scatter in each bin
inhibit fitting a reliable function that predicts $f_{{\rm bar},z=0.3}$.
Table~\ref{tbl:ferengi_bar} summarizes the correction types for the
\ferengi{} data; 71\% of the galaxies simulated did not meet the GZH
requirements for well-sampled vote fractions for $f_\mathrm{bar}$, as
compared to 2\% for $f_\mathrm{features}$. Similar results were obtained
for all higher-order Tasks in GZH (Section~\ref{ssec:higher_order_tasks}).

>>>>

The technique for redshift debiasing (Section~\ref{sec:debiasing}) requires
consistent measurements of the galaxy surface brightness. For both the
\redshifted{} calibration images and the \hst{} images which have their
morphologies corrected, we calculate the mean surface brightness $\mu$ within the
effective radius ($R_e$) as:

@@

Figure~\ref{fig:zeta_examples} shows the change in vote fraction and the
best-fit model for a random set of galaxies in the \ferengi{} sample. The
results show that there is a clear dependence on redshift for the observed
changes in \ffeatures. To examine whether a better fit should involve
additional parameters, we examine the global dependence of $\zeta$ on the
galaxy surface brightness. This is motivated by the fact that brighter galaxies
would presumably be easier to identify; \citet{bam09} and \citet{wil13} found
changes in \ffeatures{} depend on a combination of physical size and
absolute magnitude. To simplify our model, we subsume these into a single
parameter $\mu$.

@@

\begin{figure}
\center
\includegraphics[width=0.5\textwidth]{figures/zeta_mu.pdf}
\caption{All fits for the \ferengi{} galaxies of the vote fraction dropoff
parameter $\zeta$ for \ffeatures{} as a function of surface brightness. This
includes only galaxies with a bounded range on the dropoff
($-10<\log(\zeta)<10$) and sufficient points to fit each function (37~original
galaxies, each artificially redshifted in eight bins over a range from $0.3\lesssim z_\mathrm{sim}\lesssim1.0$).}
\label{fig:zeta_mu}
\end{figure}

@@

The criteria for determining whether a region of this space is single-valued,
and therefore correctable, is as follows: In each surface brightness and
redshift bin, the relationship between \ffeaturesz{} and \ffeaturesrest{} is
modeled by fitting the data with polynomials of degrees $n=3$, 2, and 1, and
using the best formal fit out of the three as measured by the sum of the
residuals. These fits are shown as the dashed black lines in
Figure~\ref{fig:f_vs_f}. Flat regions of the bins are areas in which there is
not a clear single-valued relationship between \ffeaturesz{} and
\ffeaturesrest. We quantify this by measuring the slope of the best-fit
polynomial; regions of the bins with a slope less than 0.4 are considered
\emph{not} one-to-one, and therefore \ffeaturesz{} cannot be boosted to its
\ffeaturesrest{} value.  Galaxies in this region are referred to as the
\emph{lower limit} sample, because the most stringent correction available is
that the weighted \ffeatures{} is a lower limit to the true value.  These
regions are highlighted in blue in Figure~\ref{fig:f_vs_f}. Uncolored (white)
regions of the plot have sufficiently high slopes to consider the relationship
as single-valued; galaxies in these regions are considered ``correctable'', and
only these are used in measuring the parameters for the $\zeta$ function
(Section~\ref{ssec:zeta}). Only surface brightness/redshift bins with at least
5~galaxies were considered; regions with fewer than 5~galaxies are considered
to have ``not enough information'' to determine the \ffeaturesz{} and
\ffeaturesrest{} relationship, colored gray in Figure~\ref{fig:f_vs_f}. These
galaxies are hereafter referred to as the ``NEI'' sample.

@@

For the ``lower-limit'' galaxies for which a single debiased \ffeatures{} value
cannot be confidently assigned, the \emph{range} of debiased values is
estimated using a method visualized in Figure~\ref{fig:f_vs_f_zoom}. This uses the
\ferengi{} simulated data to analyze the range of intrinsic \ffeaturesrest{}
values for any given observed \ffeatures{} value, again as a function of
surface brightness and redshift. In each unclassifiable (shaded blue) $z$,$\mu$ bin, the spread of
intrinsic values of \ffeaturesrest{} for four ranges of observed \ffeatures{} is computed.
The range of intrinsic values is defined as the inner 80\% of the data,
represented by the orange bars in Figure~\ref{fig:f_vs_f_zoom}. For any galaxy which
cannot be directly debiased by the $\zeta$ method, these ranges are used to
denote the upper and lower limits on the expected values \ffeaturesrest{} as a
function of the observed \ffeatures.

@@

Few galaxies in the sample have sufficiently high corrections to completely
change them from being confidently ``smooth'' to ``featured'' following the
correction for redshift bias (Figure~\ref{fig:debiased_corrections}). As a
check, highly boosted galaxies are compared to the expert classifications in
CANDELS \citep{kar15}.  There are only eight galaxies that were strongly
boosted (\ffeatures$<0.2$ and \fbest$>0.5$) in GZH and also appear in the
CANDELS expert sample. Of those eight galaxies, \citet{kar15} classify 5 as
spheroids, 2 as disks, and 1 as irregular/disky. The \fbest{} values for GZH
are all between 0.5 and 0.6, making them intermediate disk candidates. One of
the spheroids has a hint of extended emission to the south that may have been
missed by the CANDELS experts, but almost all images are dim and with very low
surface brightnesses in the ACS imaging, making positive identification
challenging. Since the surveys use different rest-frame filters and the overlap
sample is tiny, though, detailed comparisons between the overall morphologies
are highly difficult.

@@

Each image is listed under a unique project ID (eg,~AHZ000001); the actual
galaxy in the image is identified by the combination of the OBJNO and original
survey. For each of the 55~responses in the GZH decision tree, the following
classification data is provided: for each question, $\tt N_{votes}$ is the
number of classifiers who answered that question. For each unique answer, $\tt fraction$
is the fraction of classifiers who selected that answer ($\rm N_{answer}/N_{votes}$), and
$\tt weighted$ is the weighted fraction, which takes into account overall
consistency (Section~\ref{ssec:weighting}).

@@

The comparison and analysis of the GZH morphologies is done on a set of matched,
volume-limited samples at $z<1.0$ and $m_{I|i|z} < 22.5$~mag.  The redshift and
magnitude limits for the sample are chosen to match the shared constraints for
the shallowest depths \citep[GEMS;][]{bun05} and the limits on morphological
reliability \citep[COSMOS/ZEST;][]{sca07} in the comparison catalogs. We match
all catalogs against GZH using a positional radius of 0.5\arcsec.

@@

Figure~\ref{fig:comparisons_odd} analyzes the same sets of matched galaxies, but
compares the data to the \fodd{} parameter in GZH to examine its correlation with
irregular morphologies. The relationship in this case is less clear; for the
COSMOS catalogs, in particular, early-types, spirals, and irregulars are
roughly equally probable at \fodd$>0.5$. In both AEGIS and GOODS (there is no
measurment of irregular structure for GEMS in \citealt{hau07}), the proportion
of irregular galaxies does increase with \fodd. However, the actual number of
galaxies at \fodd$>0.5$ is quite small ($N\lesssim100$) in both samples, and so
estimating an accurate fraction is less precise.  We also note that this
particular question ({\it ``anything odd?''}) has somewhat uncertain phrasing
in GZH, and that the design of the tree explicitly separates clumpy from
irregular or merging morphologies, which would have similar properties in
automatic techniques. A more accurate selection of irregular morphologies in
GZH should use a weighted combination of \fbest, \fclumpy, and \fodd.

@@

To determine the rate at which the vote fraction for any Task decreases with
redshift or surface brightness, data from simulated \ferengi{} images were
modeled with functions in discrete redshift and surface brightness bins
(Section~\ref{ssec:zeta_results}). A cut of $N>10$ was placed on the number of
votes to reduce the error in the vote fractions and ensure they were
well-sampled before applying a fit to the data. This requirement was only met
for Task~01, which significantly reduced the goodness-of-fit for the model
functions as compared to the smooth/features task. Here we show an
example of results obtained for the bar Task. Figure~\ref{fig:f_vs_fbar}
shows that both the low sample rate and the degree of scatter in each bin
inhibit fitting a reliable function that predicts $f_{{\rm bar},z=0.3}$.
Table~\ref{tbl:ferengi_bar} summarizes the correction types for the
\ferengi{} data; 71\% of the galaxies simulated did not meet the GZH
requirements for well-sampled vote fractions for $f_\mathrm{bar}$, as
compared to 2\% for $f_\mathrm{features}$. Similar results were obtained
for all higher-order Tasks in GZH (Section~\ref{ssec:higher_order_tasks}).
